\newcommand{\zzket}{\braket{00}}
\newcommand{\zoket}{\braket{01}}
\newcommand{\ozket}{\braket{10}}
\newcommand{\ooket}{\braket{11}}

\newcommand{\zzzket}{\braket{000}}
\newcommand{\zzoket}{\braket{001}}
\newcommand{\zozket}{\braket{010}}
\newcommand{\zooket}{\braket{011}}
\newcommand{\ozzket}{\braket{100}}
\newcommand{\ozoket}{\braket{101}}
\newcommand{\oozket}{\braket{110}}
\newcommand{\oooket}{\braket{111}}

\section{Quantum Teleportation}

\begin{remark}
  Say I want to communicate my qubit $q$ with you, but I don't have a
  quantum channel to you. I do have a classical channel on which I can
  sends ones and zeroes.

  If my qubit $q$ were in a basis state, everything would be easy (I'd
  just send a zero or one). But the point is that it might not be. It
  might even be in a state I don't know!

  If we share an EPR pair already, here's what we can do. First, let the
  quantum state be notated:

  \begin{nedqn}
    \parens{
      \alpha \zket + \beta \oket
    }
    \otimes
    \sqtot
    \parens{
      \zzket + \ooket
    }
  \end{nedqn}

  \noindent
  (The first qubit is the one I wish to teleport, the second is my half
  of the EPR pair, and the third is your half of the EPR pair).

  First, I will apply a CX of my qubit with my half of the EPR pair.
  This gives us:

  \begin{nedqn}
    \sqtot
    \parens{
      \alpha \zzzket
      + \alpha \zooket
      + \beta \oozket
      + \beta \ozoket
    }
  \end{nedqn}
\end{remark}

\begin{remark}
  Now imagine I measure \emph{my} half of the EPR pair. If I measure
  zero, then the reduced state is:

  \begin{nedqn}
    \alpha \braket{000} + \beta \braket{101}
  \end{nedqn}

  \noindent
  Now, if my friend measures the third qubit, they will get a zero/one
  result with the correct probabilities $\norm{\alpha}^2,
  \norm{\beta}^2$. However, they have not truly been sent my qubit. The
  reason is that they cannot perform quantum operations on it with the
  expected result.

  For instance: say I produced $\alpha \zket + \beta \oket$ by Hadamard
  operation on $\zket$. That is, $\alpha = \beta = \sqtot$. In that
  case, after teleportation, I should be able to tell my friend to apply
  $\mtxH$ to recover $\zket$.

  But this won't work for them. The problem is the first qubit of state
  that \emph{I} hold onto. I have entangled the first qubit with the
  third. So when my friend applies $\mtxH$ to their third qubit, they
  are performing:

  \begin{nedqn}
    \parens{\mtxI \otimes \mtxI \otimes \mtxH}
    \parens{\sqtot \zzzket + \sqtot \ozoket}
  \eqcol
    \parens{
      \sqtot \zzket
      \otimes
      \mtxH \zket
    }
    +
    \parens{
      \sqtot \ozket
      \otimes
      \mtxH \oket
    }
  \\
  \eqcol
    \half
    \parens{\zzzket + \zzoket}
    +
    \half
    \parens{\ozzket - \ozoket}
  \end{nedqn}

  \noindent
  If only the states $\zzzket, \ozzket$ could \emph{constructively}
  interfere! If only the states $\zzoket, \ozoket$ could
  \emph{destructively} interfere!
\end{remark}

\begin{remark}
  We can fix the problem by applying $\mtxH$ to the first qubit. This
  gives us:

  \begin{nedqn}
    \parens{
      \mtxH \otimes \mtxI \otimes \mtxI
    }
    \parens{
      \alpha \zzzket + \beta \ozoket
    }
  \eqcol
    \alpha \sqtot \zzzket
    + \alpha \sqtot \ozzket
    + \beta \sqtot \zzoket
    - \beta \sqtot \ozoket
  \end{nedqn}

  We may now measure the first qubit. If we get zero, we know that the
  final state is $\alpha \zzzket + \beta \zzoket$. We can send $00$ to
  our friend so that they know that their qubit is good.

  If we read one, then $\alpha \ozzket - \beta \ozoket$. We should send
  $01$ to our friend so they know to correct their qubit by a $\mtxZ$
  phase-flip gate.
\end{remark}

\begin{remark}
  What if we had read a one first? Then our state would be $\alpha
  \zooket + \beta \oozket$. We can send an initial one so that they do a
  $\mtxX$ not operation on their qubit. That gives: $\alpha \zozket +
  \beta \oooket$. Now we are in the same scenario as before.
\end{remark}

\begin{remark}
  Note: there is no faster-than-light communication here. Without
  sending the two classical bits, the other person does not know how to
  correct their qubit. Also: we have not cloned our state -- we've
  actually \emph{disentangled} the EPR pair.
\end{remark}
