\documentclass[11pt, oneside]{amsart}

\newcommand{\calC}{\mathcal{C}}
\newcommand{\opD}{\operatorname{D}}
\newcommand{\opDiv}{\operatorname{div}}

\usepackage{geometry}
\geometry{letterpaper}

\usepackage{ned-common}
\usepackage{ned-calculus}
\usepackage{ned-linear-algebra}
\usepackage{ned-stats}

\begin{document}

\title{Div, Grad, Curl}
\maketitle

\section{Real-Valued Derivative, Partials, and Gradient}

\subsection{Derivative}
\begin{enumerate}

\item With a function of a single variable, we have already defined the
\emph{derivative}:

\begin{nedqn}
  \fderiv{x} f(x)
\eqcol
  \lim_{a \to x}
  \frac{
    f(a) - f(x)
  }{
    a - x
  }
\\\eqcol
  \lim_{t \to 0}
  \frac{
    f(x + t) - f(x)
  }{
    t
  }
\end{nedqn}

\item The notation $\fderiv{x} f(x)$ is called \define{Leibniz
notation}. There is also \define{Lagrange notation}, in which we write
$f'(x)$, which is common in univariate settings, but it will be
difficult to know what is being differentiated in multivariate settings.
There is \define{Newton notation}, where we write $\dot f$; this is used
in physics sometimes, the presumption is always that we are
differentiating against time. Last, there is \define{Euler notation},
where we write $\opD_x f(x)$.
\end{enumerate}

\subsection{Partial Derivative, Directional Derivative}

\begin{enumerate}
\item Let's begin to generalize the univariate derivative to a function
of many variables. For now, we'll keep $f$ mapping to a single real
variable. Then we define the \define{partial derivative}:

\begin{nedqn}
  \fpartial{x_i} f(x_1, \ldots, x_n)
\eqcol
  \lim_{t \to 0}
  \frac{
    f(x_1, \ldots, x_i + t, \ldots, x_n)
    -
    f(x_1, \ldots, x_n)
  }{
    t
  }
\end{nedqn}

\item A very typical case is going to be when, instead of treating
separate parameters, we can consider the parameters to be packaged as a
``vector'', and the domain of the function is a normed vector space. A
very typical case will be $f: \R^n \to \R$. In that case:

\begin{nedqn}
  \fpartial{x_i} f(\vx)
\eqcol
  \lim_{a \to \vx_i}
  \frac{
    \ff{\tilde\vx} - \ff{\vx}
  }{
    \norm{
      \tilde\vx - \vx_i
    }
  }
\intertext{where}
  \tilde{\vx}
\eqcol
  \parens{\vx_1, \ldots,a, \ldots, \vx_n}
\intertext{we can also write}
  \fpartial{x_i} f(\vx)
\eqcol
  \left.
    \fderiv{t} f(\vx + t\ve_i)
  \right|_{t=0}
\intertext{where $\ve_i$ is the $i$-th basis vector.}
\end{nedqn}

\item The partial derivative effectively restricts $f$ to a function of
just $x_i$, where all other parameter values $x_j$ are held constant as
$\vx_j$. Then, we calculate the typical univariate derivative at $x_i$.
This partial derivative is basically the slope of the function in the
direction of $x_i$.

\item Sometimes people write $f_x$ or $f'_x$ for the partial derivative.
The first looks like a restriction of the function $f$ to the
$x$-variable, instead of a derivative; I wouldn't use that. $f'_x(x, y)$
is similar to the Lagrange notation, but I like the Leibniz notation
better ($\fpartial{x}$). The Euler notation $\opD_x$ is also good.

\item We typically say ``partial derivative'' when we are talking about
a derivative with respect to either an individual ``parameter'' or a
``coordinate'' basis vector. But that might be a bit arbitrary of us. If
$f$ maps from a vector space, we can take a derivative along \emph{any}
direction in the space. We call this a \define{directional derivative}:

\begin{nedqn}
  \opD_{\vv} f(\vx)
\eqcol
  \lim_{t\to 0}
    \frac{
      f(\vx + t\vv) - f(\vx)
    }{
      t
    }
\intertext{equivalently}
\eqcol
  \left.
    \fpartial{t} f(\vx + t\vv)
  \right|_{t=0}
\end{nedqn}

\item Wikipedia at least writes $\grad_{\vv} f(\vx)$, but both that
notation (and $\opD_{\vv}$, tbh) are a bit odd because they seem to
suggest the existence/use of \emph{other} derivatives. The notation
$\fpartial{\vv} f(x)$ attempts to generalize the partial derivative
notation, but ChatGPT suggests this may not be a standard notation. I do
think maybe the $\opD_{\vv}$ notation is probably the most preferred.
One downside is that, if $f$ is not (total) differentiable at $\vx$, use
of $\opD_{\vv} f(\vx)$ might be misleading if it suggests the total
derivative $\opD f(\vx)$ exists.
\end{enumerate}

\subsection{Total Derivative}
\begin{enumerate}
\item A function $g$ is a \define{first-order approximation} of $f$ if
at $\vx$ if, for any direction $\vu$:

\begin{nedqn}
  \lim_{t \to 0}
  \frac{
    g(\vx + t\vu) - f(\vx + t\vu)
  }{
    t \norm{\vu}
  }
\eqcol
  0
\end{nedqn}

\item The function $g(\vu) = f(\vx) + \opD_{\vu} f(\vx)$ is a
first-order approximation of $f$ at $\vx$, almost by definition. Of
course, this first-order approximation only exists if $\opD_{\vu}$
exists.

However, if $\opD_{\vu} f(\vx)$ is too ``weird'' (directional derivative
changes too wildly as direction changes), we will not call $f$
\define{(total) differentiable} at $\vx$. For $f$ to be considered
(total) differentiable at $\vx$, we will require that there exists an
\define{affine} first-order approximation. This occurs exactly when
$\opD_{\vu} f(\vx)$, when treated as a function of $\vu$, is
\emph{linear}. At this point, we define $\opD f(\vx)$ to be the linear
functional:

\begin{nedqn}
  \left(\opD f (\vx)\right) \vu
\defeq
  \opD_{\vu} f(\vx)
\end{nedqn}

\noindent
We call $\opD f(\vx)$ the \define{(total) derivative} of $f$ at $\vx$.
It is like the ``slope'' of the (multivariate) affine first-order
approximation for $f$ at $\vx$.

\item In the univariate case, there was no need to separate the ideas of
partial/directional derivative and total derivative. In the multivariate
case, the existence of a directional derivative along every direction
does not imply that a \emph{linear} total derivative exists. The
linearity of directional derivatives is exactly what is needed
for a first-order approximation of $f$ at $\vx$ to exist.
\end{enumerate}

\subsection{Gradient}
\begin{enumerate}
\item We can define the \define{gradient}. It is a vector, and to write
it down, we need to have fixed a coordinate basis. The gradient is
defined:

\begin{nedqn}
  \gradient f(\vx)
\eqcol
  \begin{pmatrix}
    \fpartial{x_1} \ff{\vx}
    \\
    \vdots
    \\
    \fpartial{x_n} \ff{\vx}
  \end{pmatrix}
\end{nedqn}

\noindent
The gradient is useful because (when $f$ is (total) differentiable):

\begin{nedqn}
  \left(\opD f(\vx) \right) \vu
=
  \opD_{\vu} f(\vx)
=
  \gradient f(\vx) \cdot \vu
\end{nedqn}

\noindent
Provided the function is relatively well-behaved, the gradient allows us
to calculate the directional derivative in any direction. Because $\opD
f(\vx)$ is required to be linear, its value at $n$ basis vectors is
enough to specify it.

\item Sometimes the $\gradient$ symbol is treated as a column of partial
derivative operators:

\begin{nedqn}
  \gradient
\eqcol
  \begin{pmatrix}
    \fpartial{x_1}
    \\
    \vdots
    \\
    \fpartial{x_n}
  \end{pmatrix}
\end{nedqn}

\noindent
In which case, $\gradient f(\vx)$ is treated as if it were a vector
($\gradient$) times a scalar: $f(\vx)$. We'll see other differential
operators (divergence and curl) are defined with notation that is based
on this interpretation of the $\gradient$ symbol. Still, I think the
notation is really secondary to the underlying concepts/definitions. We
should try to understand things in terms of the underlying concepts, not
the notation.

\item Note that our gradient has the shape of a ``column'' vector in the
space, so the gradient is often interpreted as a \emph{direction}
(``directly uphill''). On the other hand, $\opD f(\vx)$ is a linear
functional (a map from vectors to scalars), and so can be interpreted as
a \emph{co-vector} (a ``row'' vector). In a sense, $\opD$ and
$\gradient$ are the same, just \emph{transposes} of each other.

% TODO: produce a valid counterexample.

% \item However, we should be careful to note that this only true for
% somewhat well-behaved functions! It's not always true that you can
% properly calculate a directional derivative simply by knowing the
% gradient along some axes. Take a silly example:

% \begin{nedqn}
%   f(x, y)
% \eqcol
%   \begin{cases}
%     x \text{ if } y \ne x \\
%     0 \text{ if } y = x
%   \end{cases}
% \end{nedqn}

% \noindent
% Here, $\fpartial{x} f(0, 0) = 1$, $\fpartial{y} f(0, 0) = 0$. But the
% derivative along the direction defined $y = x$ is 0, and not
% $\frac{\sqrt{2}}{2}$. You can even create smooth examples!

% TODO: this example may be wrong?

\end{enumerate}

\section{Vector Valued Functions}

\begin{enumerate}

\item The other direction of generalization is to consider
\define{vector-valued functions}. A vector-valued function $\vec{f}:
\R^m \to \R^n$ is also sometimes called a \define{vector field}. A
vector space is just a \emph{space} of vectors, like $\R^n$. ``Field''
here is at odds with the definition in \emph{algebra} (where the
rational numbers, $\mathbb{Q}$, is an example of an algebraic field).
The use of vector field as an assignment of a vector to each point in a
vector space is really more of a physics and vector calculus name.

\item Famous examples of vector fields are electric fields, magnetic
fields, gravitational fields, fluid velocity fields.

\item Our prior definition of a partial derivative (and of a directional
derivative) will still apply. We just need to note that the numerator in
those definition is a vector.

We will need the vector space to be \emph{normed}, because we need a
notion of convergence so that a limit can be defined.

\item We will also generalize the total derivative. We previously had
that $\opD f(\vx)$ was the linear transformation that mapped a vector to
the directional derivative of $f$ at $\vx$ along that vector.
$\left(\opD \mtx{F}(\vx)\right)(\vu)$ will remain the same: except now
the directional derivative of $\mtx{F}$ at $\vx$ along $\vu$ is
\emph{vector-valued}.

\item In the scalar-valued case because $\opD f(\vx)$ is a linear
functional, we could represent it as a row vector. The vector-valued
total derivative will correspond to a \emph{matrix} that we call the
\define{Jacobian}:

\begin{nedqn}
  \mJ_{\mtx{F}}(\vx)
\eqcol
  \begin{bmatrix}
    \grad\tran \mtx{F}(\vx)_1 \\
    \grad\tran \mtx{F}(\vx)_2 \\
    \vdots \\
    \grad\tran \mtx{F}(\vx)_n \\
  \end{bmatrix}
\end{nedqn}

\noindent
Here, $\mtx{F}(\vx)_i$ just means the restriction of the value of
$\mtx{F}$ to the $i$th-coordinate. $\gradient\tran \mtx{F}(\vx)_i$ is
effectively the representation of the linear functional
$\opD_{\mtx{F}_i}(\vx)$.

For $\mtxF : \R^m \to \R^n$, $\mJ_{\mtx{F}}(\vx)$ is an $n$-by-$m$
matrix. It maps vectors of $\R^m$ to the vector-valued directional
derivative along that vector.

\item As before, provided the function really is differentiable and thus
the total derivative exists, then:

\begin{nedqn}
  (\opD \vec{f} (\vx)) \vv
\eqcol
  \mJ(\vx) \vv
\end{nedqn}

\noindent
Note that the Jacobian is more a proper representation of the total
derivative than the gradient is, in the sense that you don't have to
transpose it.

\item We don't really define an analogue of the gradient for
vector-valued functions.

\end{enumerate}

\section{Line and Surface Integrals}

\subsection{Line Integral}

\begin{enumerate}

\item We can next define a \define{line integral} (also called a path
integral). First we have to define a \define{curve}. A curve $\calC$ is
a subset of a vector space (usually $\R^n$) with the following
requirement: there is a continuous map $r: [a, b] \to \R^n$, and the
image of $r$ is $\calC$ ($[a, b]$ is a closed interval in $\R$, of
course). Everything probably remains fine whether the curve is closed or
open.

\item Now we can define the line integral of $f$ along a curve $\calC$
defined by $r$. We'll want $r$ to be piecewise differentiable. It is
defined to be:

\begin{nedqn}
  \int_\calC
  f \diff{s}
\eqcol
  \int_a^b
  \ff{\vr(t)}
  \norm{\vr'(t)}
  \dt
\end{nedqn}

\noindent
Here the notation $\vr(t)$ emphasizes that this is a vector-valued
function of $t$. Note that the ``rate'' at which we move through the
curve (which is arbitrary) is corrected by $\norm{\vr'(t)}$. This seems
to make the line integral independent on the parameterization of
$\calC$, but in fact any parameterization will give the same integral
result (would need proof!).

\item We define the \define{length} of a curve as $\int_\calC 1
\diff{s}$.

\item For now, I will consider $\int_\calC f \diff{s}$ as pure
notation. It feels pretty meaningless and arbitrary at present.

\item Note that the path-integral concept trivially generalizes to a
vector-valued function $\mtx{F}$.

\item It feels a little bit like length should be a property of the
curve in the vector space itself and a definition of unit length for
unit vectors. It feels like a bit of an intrusion that we would need to
introduce to a parameterization like $\vr$, and use it so centrally in
the definition of the length of a curve.

\item But I believe differentiability of the curve will always be
necessary because length of the curve has to do with a integral of the
magnitude of the tangency vector. And you will want to control the rate
at which you trace the curve, so that you don't ``cheat'' by moving
through different parts at different rates.

I can't see a cleaner, simpler way to do this than by introducing $\vr$.
Whatever we do probably will always end up equivalent to introducing
$\vr$.

\end{enumerate}

\subsection{Surface Integral}

\begin{enumerate}

\item We can define a corresponding \define{surface integral}. We must
again first define what we mean by a \define{surface}. A surface $S$ is
a subset of a vector space (typically $\R^n$) with the following
property: there exists a continuous mapping $r: U \to \R^n$ ($U$ should
be an open set in $\R^2$) where $S$ is the image of $r$.

I think that it's fine whether the set is open or closed, so long as
it's not too misbehaved. It would probably be best if the
parameterization is bijective, but that probably isn't strictly
necessary.

\item We should next define the notion of \define{surface area} (just as
we defined curve length). My first attempt was to try define this as:

\begin{nedqn}
  \iint_S
  1 \diff{S}
& \stackrel{?}{:=} &
  \iint_U
  \det \left(\mtxJ_{\vr}(\vu)\right)
  \diff{\vu}
\end{nedqn}

\noindent
Here we use that $\det \mtxJ_{\vr}(\vu)$ is the amount by which area is
stretched by $\vr$ in the vicinity of $\vu$. However, since $\det$ is
only defined for \emph{square} matrices, this won't quite work when we
are mapping from two-dimensions to three!

\item Really, we should use the cross-product:

\begin{nedqn}
  \iint_S
  1 \diff{S}
\defeqcol
  \iint_T
  \norm{
    \frac{\partial \vr}{\partial x}(\vu)
    \cross
    \frac{\partial \vr}{\partial y}(\vu)
  }
  \diff{x} \diff{y}
\end{nedqn}

\noindent
This is correct, because $\va \cross \vb$ has magnitude equal to the
area spanned by the two vectors: $\va \cross \vb = \norm{\va}\norm{\vb}
\sin(\theta)$.

\item We can match this to my intuition to use the determinant. First,
note:

\begin{nedqn}
  \det A
\eqcol
  ad - bc
\intertext{and}
  \colvec{a \\ c \\ 0} \cross \colvec{b \\ d \\ 0}
\eqcol
  \parens{
    ad - bc
  } \vk
\end{nedqn}

\noindent
So while the correct definition is definitely to use the cross-product,
the determinant idea was on the right track.

\item I would note that area is defined by height times width, where
those are measured orthogonally. We're assuming that $x$ and $y$ are
orthogonal directions in the space. So the local stretching of a space
depends on (1) how much differences in $x$ and $y$ get stretched, and
(2) how much they remain orthogonal in the target space.

\item Our definition in terms of cross-product of partial derivatives
kind of only works if $\mtxF$ has values in $\R^3$. But you can probably
use some variation of my determinant idea for hypersurfaces of dimension
in $n-1$ in spaces of $n>3$ dimensions.

\end{enumerate}

\section{Flux and Divergence}

\subsection{Flux}
\begin{enumerate}

\item For a field $\mtxF: \R^n\to\R^n$ (typically $n=3$), we define the
\define{flux} $\Phi_{\mtxF}(S)$ of $\mtxF$ through the surface $S$ to
be:

\begin{nedqn}
  \Phi_{\mtxF}(S)
\defeqcol
  \iint_S
  \mtxF(\vu) \cdot \vn(\vu)
  \diff{\vu}
\end{nedqn}

\noindent
What are we integrating? $\mtxF(\vu) \cdot \vn(\vu)$. What is
$\vn(\vu)$? It will be the unit-vector that is \define{normal} to the
surface $S(V)$ at position $\vu$. Sometimes we call $\mtxF(\vu) \cdot
\vn(\vu)$ the \define{flux density} of $\mtxF$ at $\vu$.

\item We sometimes write $\Phi_{\mtxF}(S) \defeq \iint_S \mtxF \cdot
\diff{\mtx{S}}$.

\item Examples are fluid flow through a surface (a mass flux), electric
flux, magnetic flux, heat flux (amount of heat transfer across a
surface).

\item The idea is, this is the amount of transfer across a surface. We
need to take the inner product of $\mtxF$ with $\vn$ to ignore the
component which is directed \emph{parallel} to the surface, so that we
can isolate the component \emph{across} the surface. The flux is scalar
valued, because it is the \emph{amount} of $\mtxF$ directed across the
surface $S$, which doesn't possess a single ``direction.''

\end{enumerate}

\subsection{Divergence}
\begin{enumerate}

\item The flux is not measured at a single point. It is measured across
a surface. However, we can define the \define{divergence} of a vector
field $\mtxF$ at a point $\vu$:

\begin{nedqn}
  \opDiv \mtxF(\vu)
\defeqcol
  \lim_{\operatorname{Vol} V_i \to 0}
  \frac{
    \Phi_{\mtxF}(V_i)
  }{
    \operatorname{Vol} V_i
  }
\end{nedqn}

\noindent
$V_i$ is a series of volumes about $\vu$ with volume $\operatorname{Vol}
V_i$ converging to zero. For instance, we could consider
$\epsilon$-balls about $\vu$ with radius falling to zero:

\begin{nedqn}
  \opDiv \mtxF(\vu)
\defeqcol
  \lim_{\epsilon \to 0}
  \frac{1}{\operatorname{Vol} B_\epsilon(\vu)}
  \iint_{S_\epsilon(\vu)}
  \mtxF(\vx) \cdot \vn(\vx)
  \diff{\vx}
\end{nedqn}

\noindent
Here, $S_\epsilon(\vu)$ is the $\epsilon$-sphere centered at $\vu$.

\item It might appear slightly odd at first that we are dividing by the
\emph{volume} of the ball, while we are integrating across a
two-dimensional surface. But I believe this works out if we want the
divergence to be an infinitesimal \emph{flux density}. That is, the
amount of flux \emph{per unit area} of an infinitesimal ball centered at
$\vu$. We need to divide by volume because we need to divide by surface
area (to get a flux density) times radius (to make this an
infinitesimal).

\item Soon we will prove that, if $\mtxF$ is differentiable at $\vu$,
then:

\begin{nedqn}
  \opDiv \mtx(\vu)
\eqcol
  \sum_i \fpartial{x_i} \mtxF(\vx) \cdot \ve_i
=
  \sum_i \fpartial{x_i} F_i(\vx)
\end{nedqn}

\noindent
Where $F_i(\vx)$ is the $i$-th coordinate of $\mtx{(\vx)}$.

Note that if the equality holds, then the summation must be constant
regardless choice of basis. The obvious requirement is that the basis
must be orthonormal.

\item If you treat the nabla operator $\nabla$ as a vector of partial
operators, then you have:

\begin{nedqn}
  \gradient \cdot \mtxF
\eqcol
  \sum_i \fpartial{x_i} F_i
= \opDiv \mtxF
\end{nedqn}

\noindent
You don't have to take this \emph{too} literally. But this is a common
way to notate $\opDiv \mtxF$.

\item \TODO{Prove that divergence is sum of partials.}
\item \TODO{Notation for integration; what is the differential symbol
for}?
\item \TODO{What does it mean to do a dot product with an infinitesimal
in an integral? What is that notation? Used in flux formula, for
instance.}
\item \TODO{Define curl}.
\item \TODO{Show nabla cross product notation for curl}.
\item \TODO{Laplacian?}
\item \TODO{Line integral independent of path in conservative fields}

\end{enumerate}

\end{document}
