\subsection{Some Properties of CoM}

\begin{enumerate}
  \item The center of mass is like the average position of the
  \emph{grams} in the object. This is what it means to be weighted by
  mass.

  \item The CoM might not be in the material of the object, but if the
  object is bounded, of course the CoM must lie within those bounds.

  \item The center of mass will always lie within any planes of symmetry
  that exist for the object.

  \item For every plane that passes through the center-of-mass, the sum
  $\sum_i m_i \vx_i$ of every subpart on one side of the plane is
  exactly the same as the sum for everything on the other side of the
  plane, excepting a negative sign.

  \item If you project all points $\vx_i$ into a subspace that includes
  the center of mass, then the mass-weighted sum of the projections is
  still equal to center of mass. If the subspace does not include the
  center of mass, then the center of mass within the subspace lies at
  the projection of the center of mass into the subspace.

  \item Consider subobjects A and B, with masses $M_A, M_B$. Now combine
  the objects. The center of mass is $\frac{M_A}{M} \vx_A +
  \frac{M_B}{M} \vx_B$, where $M = M_A + M_B$. This is because:

  \begin{nedqn}
    \frac{1}{M} \sum_i m_i \vx_i
  \eqcol
    \frac{1}{M}
    \parens{
      \sum_{i \in A} m_i \vx_i
      +
      \sum_{i \in B} m_i \vx_i
    }
  \\
  \eqcol
    \frac{1}{M}
    \parens{
      M_A \vx_A + M_B \vx_B
    }
  \end{nedqn}

  \item Feynman then goes on a long philosophical ramble about how
  Newton's laws ``reproduce'' themselves on greater and greater levels.
  That is: if Newton's laws apply to individual particles, they also
  apply to aggregates. Feynman notes that these are the kind of laws
  that we are most likely to find first, because they occur at larger
  and thus observable scales. He notes that the true laws at atomic
  level are not exactly Newton's, but when aggregates are studied,
  Newton's are a good approximation (and get better as we aggregate
  further).
\end{enumerate}
